---
title: "1) Species relationship with suitability - data based approach"

output:
  html_document:
    #keep_md: true
    code_folding: "hide"
---

## Analysis

**Purpose:** To quantify the distribution of simulated species response (occurence & abundance) in relation to habitat suitability.

Each simulated habitat suitability map was classified into 20 bins (0-5,5-10...). Zonal statistics were performed using the classified suitability rasters and simulated abundance to calculate the total abunance within each suitability class. The simulated abunance rasters were converted to binary occurence rasters and the total number of occupied cells within each suitability class was calculated using zonal statistics again. These were then converted to be proportional to the total number of cells within each suitability class.
```{r  }
library(dplyr)
library(ggplot2)
library(ggpubr)
setwd("D:/SyncProjects/PhD/Chapters/RQ1/Submission/SupplementaryMaterial/Supplementary_4")

prevalence <- read.csv("data/prevalence.csv")
landscape75 <- read.csv("data/landscape_metrics75.csv")
landscape85 <- read.csv("data/landscape_metrics85.csv")
dat <- read.csv("data/raw_structure.csv")

dat$break_pt <- dat$break_pt*5
dat$abun_prop <- dat$abun_sum/dat$suit_freq
theme_Publication <- function(base_size=14, base_family="Arial") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.4)),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(,size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               #legend.position = "bottom",
               #legend.direction = "horizontal",
               #legend.key.size= unit(1, "cm"),
               #legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               #plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
          ))
      
}
```
### Prevalence

The prevalence of each species increases linearly with the availability of high suitability cells (>75). The CA and NullVLP species have very little noise along this trend compared to the IBM and NullLP species. The increased variability in NullLP compared to NullVLP is likely due to prevalence being influenced by suitability values lower down the gradient.

```{r message=FALSE, warning=FALSE, fig.width = 12, fig.height=8}
prev_ls <- prevalence %>% full_join(landscape85,by=c("ls"="ls"))
# ggplot(prev_ls[prev_ls$sim==2,],aes(y=prevalence,x=total_area/160000))+
#   geom_point()+
#   theme(axis.text.x = element_text(angle = 90))+
#   ylab("Species prevalence") + xlab("Prevalence of high suitability (>75)")+facet_wrap(vars(species),ncol=2)
prev_ls[!(prev_ls$species %in% c('DO_Species5','DO_Species6','DO_Species7','DO_Species8')),] %>% ggplot(aes(y=prevalence,x=total_area/160000,color=species,linetype=species))+
  geom_point(aes(color=species,shape=species),size = 1)+
  geom_smooth(method='lm',se=FALSE,size=1.5)+
  scale_linetype_manual(values=c("dotted", "dotted","dotted", "dotted","solid", "solid","solid","solid","twodash", "twodash"))+
  scale_shape_manual(values=c(16,16,16,16,17,17,17,17,15,15))+
  scale_colour_manual(values = c("#D55E00", "#0072B2", "#009E73","#CC79A7","#D55E00", "#0072B2", "#009E73","#CC79A7","black","black"))+
   ylab("Species prevalence") + xlab("Prevalence of very high suitability (>85)")+theme_Publication()

```
# Response
Plots below show the mean proportion and mean abundance per pixel as a line with shaded areas representing the range between first and third quantiles.
Abundance of the null model scales linearly with habitat suitability, which is as expected given the intensity of the Poisson distribution also scales linearly with suitability. As suitability increases the probability of drawing a zero from the Poissson distribution decreases, resulting in the logarithmic curve when assessing proportion of cells with occurences.
Prevalence of ABM simulated species much lower than the Null model, with >0.5 prevalence in each suitability class achieved only at high suitabilities (0.6-0.8 depending on parameters). The low fecundity species has the lowest prevalence and abundance overall, but the difference between the ABM simulated species is significantly less at the highest suitabilities compared to the middle values. 
```{r message=FALSE, warning=FALSE, fig.width=10,fig.height=13}
m <- dat %>% 
  group_by(species,break_pt) %>%
  summarise_at(vars(pa_prop), funs(median(., na.rm=TRUE)))
mQuant <- dat %>% 
  group_by(species,break_pt) %>%
  summarise(Q1=quantile (pa_prop, probs=0.25), Q2=quantile (pa_prop, probs=0.50), Q3=quantile(pa_prop, probs=0.75))
m <- as.data.frame(m)
mQuant <- as.data.frame(mQuant)
m <- cbind(m,mQuant$Q1,mQuant$Q3)
names(m) <- c('species','break_pt','pa_prop','Q1','Q3')
p1 <- ggplot(data=m[m$species %in% c('EO_Species1','EO_Species2','DO_Species1','DO_Species2','DO_Species3','DO_Species4'),], aes(x=break_pt, y=pa_prop, group=species)) +
  geom_line(aes(color=species,linetype=species),size=1)+
  geom_ribbon(aes(ymin = Q1, ymax = Q3,fill=species),alpha=0.3,linetype='solid')+
  scale_linetype_manual(values=c("solid","solid","solid","solid","dashed","dotted"))+
  scale_colour_manual(values = c("#D55E00", "#0072B2", "#009E73","#CC79A7","black","black"))+
  scale_fill_manual(values = c("#D55E00", "#0072B2", "#009E73","#CC79A7","black","black"))+
  xlab('Habitat suitability')+
  ylab('Proportion of occurence')+ylim(0,1)+ ggtitle("CA Species")+theme_Publication()+theme_Publication()+theme(plot.title = element_text(hjust = 0.01))
p3 <- ggplot(data=m[m$species %in% c('DP_Species1','DP_Species2','DP_Species3','DP_Species4'),], aes(x=break_pt, y=pa_prop, group=species)) +
  geom_line(aes(color=species),size=1)+
  geom_ribbon(aes(ymin = Q1, ymax = Q3,fill=species),alpha=0.3,linetype='solid')+
  scale_colour_manual(values = c("#D55E00", "#0072B2", "#009E73","#CC79A7"))+
  scale_fill_manual(values = c("#D55E00", "#0072B2", "#009E73","#CC79A7"))+
  xlab('Habitat suitability')+
  ylab('Proportion of occurence')+ylim(0,1) + ggtitle("IBM Species")+theme_Publication()+theme(plot.title = element_text(hjust = 0.01))

plotout <- ggarrange(p1,p3,common.legend=FALSE,ncol=1)
plotout
```

### Between landscape variability
When we plot mean values for simulation replicates in each landscape there is virtually no variability in mean proportion of occurence for simulations with no endogenous processes (Null species). There is greater variability in the cellular automata species, with the greatest variability observed in the IBM species that included both population and dispersal processes. For the high fecundity IBM species (Species 2 & 4), the greatest variability between landscapes is along the middle portion of the suitability gradient. In the low fecundtiy species, teh greatest variabilty was in the higher regions of suitability.

```{r message=FALSE, warning=FALSE,fig.width = 10, fig.height=15}
dat2 <- dat %>% full_join(landscape75,by=c("ls"="ls"))
m <- dat2 %>% 
  group_by(species,break_pt,ls,area_mn) %>%
  summarise_at(vars(pa_prop), funs(mean(., na.rm=TRUE)))
p <- ggplot(data=m, aes(x=break_pt, y=pa_prop, group=ls)) +
  geom_line(aes(color=area_mn))+  scale_color_gradientn(colours = rainbow(5))
p + facet_wrap(vars(species),ncol=2) + xlab("Habitat suitability") + ylab("Proportion of occurence")+theme_Publication()

```


